---
title: "UNICEF_Report_on_C-Section_Births"
format: html
editor: visual
execute:
  echo: false
  warning: false
---

<hr>

# Global Trends in Cesarean Section Births

```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
df1 <- read_csv("unicef_indicator_1 (1).csv", show_col_types = FALSE)
df2 <- read_csv("unicef_indicator_2 (1).csv", show_col_types = FALSE)
meta <- read_csv("unicef_metadata (1).csv", show_col_types = FALSE)
combined <- bind_rows(df1, df2)

full_data <- combined %>%
  left_join(meta, by = c("country", "alpha_3_code"))
# Filter only C-section data
csection_data <- full_data %>%
  filter(indicator == "C-section rate - percentage of deliveries by cesarean section") %>%
  group_by(country) %>%
  filter(time_period == max(time_period)) %>%
  ungroup()

# Load and join with spatial map
world <- ne_countries(scale = "medium", returnclass = "sf")
world_data <- world %>%
  left_join(csection_data, by = c("name" = "country"))

# Plot
ggplot(world_data) +
  geom_sf(aes(fill = obs_value)) +
  scale_fill_viridis_c(option = "C", na.value = "gray90") +
  labs(title = "C-section Rates by Country (Most Recent Year)", fill = "%") +
  theme_minimal()

```

<hr>

This is a world map showing how often babies are born by C-section (Cesarean section) in different countries, based on the most recent data available. Each country is colored depending on how high or low their C-section rate is.

Countries that are bright yellow or orange have very high C-section rates — over 60% of births there happen by surgery.

Countries in pink and purple shades have lower rates, around 20% to 40%.

Places that are dark purple or almost blue have very low rates, under 20%.

On the right side, there’s a color bar that acts like a key, showing you how the colors match to percentages.

If you look closely:

South America, especially countries like Brazil, stands out in bright colors — meaning lots of babies are born by C-section there.

Africa mostly shows darker colors — meaning fewer C-sections.

Europe and parts of Asia are a mix, with some countries having moderate to high rates.

Some countries are gray, meaning there’s no available data for them.

In short, this map is like a colorful snapshot showing that C-section births are very common in some places but much rarer in others.

<hr>

```{r}
top10 <- full_data %>% 
  filter(indicator == "C-section rate - percentage of deliveries by cesarean section") %>%
  group_by(country) %>%
  filter(time_period == max(time_period)) %>%
  ungroup() %>%
  arrange(desc(obs_value)) %>%
  slice_head(n = 10)

# Check that the data passed to ggplot is filtered correctly
latest_data <- full_data %>%
  filter(indicator == "C-section rate - percentage of deliveries by cesarean section") %>%
  filter(time_period == max(time_period))

# Plot the data
ggplot(latest_data, aes(x = reorder(country, obs_value), y = obs_value)) + 
  geom_col(fill = "steelblue") + 
  coord_flip() + 
  labs(
    title = "C-section Rate by Country",
    x = "Country",
    y = "C-section Rate (%)"
  ) + 
  theme_minimal()
```

<hr>

This is a bar chart that shows how often babies are delivered by C-section (Cesarean section) in three countries: **Yemen**, **Afghanistan**, and **Mozambique**.

-   **Yemen** has the biggest bar (highest number).

-   **Afghanistan** comes next.

-   **Mozambique** has the smallest bar of the three, but it’s still showing a very high number.

    <hr>

```{r}
yemen_trend <- csection_data %>% filter(country == "Yemen")

ggplot(yemen_trend, aes(x = time_period, y = obs_value)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(title = "C-section Rate Over Time in Yemen", x = "Year", y = "%") +
  theme_minimal()

```

<hr>

This chart is about **C-section rates in Yemen** over time.

-   The X-axis (bottom) shows the **year** — here it’s only **2023**.

-   The Y-axis (side) shows the **C-section rate in percentage**.

-   There’s just **one small black dot** on the chart, sitting at around **10.8%**.

What this means:

We only have **one data point** for Yemen — from the year 2023.

It tells us that in 2023, **about 10.8% of babies** in Yemen were delivered by C-section.

There's no line or trend because there's no data for earlier years to connect.

<hr>

```{r}
ggplot(csection_data, aes(x = time_period, y = obs_value)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Global C-section Rate Over Time", x = "Year", y = "%") +
  theme_minimal()
```

<hr>

This chart shows how the **global C-section rate** (how often babies are born by surgery) has changed over time, from around **2008 to 2023**.

Every **small black dot** represents the C-section rate for one country in a particular year.

The **red line** shows the overall trend — kind of like the "big picture" of what's happening.

When we look at it:

The black dots are spread out — some countries have low C-section rates, some have really high ones.

The red line is **slightly going down over time**, meaning that **overall**, the average C-section rate across the world might be decreasing a little bit.

However, there are still a lot of countries with high C-section rates scattered through the years.

<hr>

<hr>

**Conclusion:**\
C-section rates are different all over the world. Some countries have very high rates, while others have very low rates. Overall, around the world, the average C-section rate has stayed mostly the same over the past few years. Good, accurate data is important because sometimes mistakes can happen and make the results look wrong.

In short, C-sections are common in some places, rare in others, and we need good information to really understand the situation.

<hr>